---
title: "SHAP Analysis in multi-class classification tasks"
author: "Ramtin Zargari Marandi"
output: html_document
---

## Load Required Libraries

We will begin by loading the necessary packages.

```{r}
library(explainer)
library(mlbench)
library(mlr3learners)
library(ranger)
library(iml)
library(dplyr)
library(ggplot2)
library(plotly)
library(cvms)
library(ggplot2)
library(gridExtra)
```

## Set Seed for Reproducibility

```{r}
seed <- 246
set.seed(seed)
```

## Load and Prepare the Data

We will use the Iris dataset for this example and define the target column (`Species`).

```{r}
# Load the Iris dataset
data("iris")
target_col <- "Species"

# Extract relevant columns from the dataset
mydata <- iris
```

## Create the Classification Task

We create a classification task using the `mlr3` package.

```{r}
maintask <- mlr3::TaskClassif$new(
  id = "my_classification_task",
  backend = mydata,
  target = target_col
)
```

## Split the Dataset

We will split the data into training and test sets.

```{r}
splits <- mlr3::partition(maintask, ratio = 0.3)
```

## Train the Model

Using the ranger algorithm, we will train a classification model.

```{r}
mylrn <- mlr3::lrn("classif.ranger", predict_type = "prob")
mylrn$train(maintask, splits$train)
```

## Make Predictions

Once the model is trained, we make predictions on the test set.

```{r}
# Make predictions on the test set
predictions <- mylrn$predict(maintask, splits$test)

# Convert predictions to a data frame
predicted_classes <- predictions$response
true_classes <- mydata[splits$test, target_col]
```

## Confusion Matrix

We create and visualize a confusion matrix using the `cvms` package.

```{r}
# Create a confusion matrix
conf_matrix <- table(True = true_classes, Predicted = predicted_classes)

# Convert the confusion matrix to a data frame with required column names
conf_matrix_df <- as.data.frame(conf_matrix)
colnames(conf_matrix_df) <- c("Target", "Prediction", "N")

# Plot the confusion matrix
plot_confusion_matrix(conf_matrix_df)
```

## SHAP (SHapley Additive exPlanations) Analysis

We will now compute SHAP values for the dataset to understand the feature importance for each class.

We now use the SHAP function to generate the SHAP plots for each class and a combined plot for all classes.

```{r}
SHAP_output <- eSHAP_plot_multiclass(
  task = maintask,
  trained_model = mylrn,
  splits = splits,
  sample.size = 30,
  seed = seed,
  subset = 0.8
)

# Display the combined SHAP plot for each class
print(SHAP_output$combined_plot)
```

## Display the combined plot for all classes together

```{r}
print(SHAP_output$combined_all_classes)
```

### Notes:
- This script loads the Iris dataset, trains a classifier using the `ranger` algorithm, makes predictions, creates a confusion matrix, and calculates SHAP values for feature importance.
- SHAP values are visualized per class and across all classes using `ggplot2`.

